---
title: "p8105_hw1_edm2152"
author: "Eric Morris"
date: "September 15, 2018"
output: github_document
---

```{r load_tidy, echo=FALSE}

library(tidyverse)

```

# Problem 1 

The code chunk below creates a data frame comprised of: 
* A random sample of size 10 from a uniform[0, 5] distribution
* A logical vector indicating whether elements of the sample are greater than 2
* A (length-10) character vector
* A (length-10) factor vector

```{r Problem_1}

problem1_df = tibble(
  random_sample = runif(10,0,5)
  greater2_logical = which(random_sample > 2)
  vector_character = character(length = 10)
  vector_factor = factor(c("low", "high", "low", "high", "low", "high", "low", "high", "low", "high"))
)

```

When I try to take the mean of each of the four variables:

The mean of the random sample variable is `r mean(random_sample)`.

The mean of the logical vector variable is `r mean(greater2_logical)`.

The mean of the character vector variable is `r mean(vector_character)`.

The mean of the factor vector variable is `r mean(vector_factor)`.

Only the mean of the random sample variable will work, because it is the only numerical variable in the data frame. 

Code chunk below that applies the as.numeric function to the 3 variables: 

```{r as_numeric, include=FALSE}

  greater2_logical_numeric = as.numeric(greater2_logical) 
  vector_character_numeric = as.numeric(vector_character)
  vector_factor_numeric = as.numeric(vector_factor)

```

Second code chunk that converts character variable from character to factor to numeric and cconverts the factor variable from factor to character to numeric.

```{r Second_Conversion}

```



# Problem 2
